---
title: "Rmarkdown-exercise-worksheet"
author: "Reid Otsuji"
date: "1/5/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Instructions:**   
Complete the worksheet using RStudio and Rmarkdown.  Some R Packages will be necessary to complete the worksheet.  
Answer all questions in this document and have fun with Rmarkdown!

**Before you begin:**
You will need to load the GGPLOT package to use the `diamonds` data set. Run the code each of the following chuncks once to install and load the GGPLOT2 package.   


**1. Install and Load the `ggplot2` package**    
(if you already have GGPLOT2 installed skip this code chunk.)
```{r, eval=FALSE, results='hide'}
# This is an example of a code chunk to install an R package
# Run this code chunk manually if you need to install the GGPLOT2 package
# click the green arrow to the left to install the package.
install.packages("ggplot2")
```

  - Run this code chunk to Load the GGPLOT2 library.  

```{r,  results='hide', warning=FALSE}
# load the GGPLOT2 library
library(ggplot2)
```

**2. In the Diamonds Data Set Description text provided below, add the proper Rmarkdown formatting to each line of text.  The `highlighted` recommendations at at the end of each line.** 

  - **optinoal:** add a link to the Diamond Data Set text below using the link formt `[link text](http://add link here)` to the dataset information.
    * The 4 C's of Diamond Quality `https://4cs.gia.edu/en-us/4cs-diamond-quality/`

---

# Prices of 50,000 round cut diamonds `H1 heading`   
## Description `H2 heading`   

**A dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:** `bold this line`

### Usage `h3 heading`

* diamonds `bullet`

### Format `h3 heading`    
**A data frame with 53940 rows and 10 variables:** `bold`    

* #### price `h4 heading, bullet`
  * price in US dollars (\$326–\$18,823) `nested bullet`   

* #### carat `h4 heading, bullet`   
  * weight of the diamond (0.2–5.01) `nested bullet`   

* #### cut `h4 heading, bullet`   
  * quality of the cut (Fair, Good, Very Good, Premium, Ideal) `nested bullet`   
* #### color `h4 heading, bullet`   
  * diamond colour, from D (best) to J (worst) `nested bullet`   

* #### clarity `h4 heading, bullet`   
  * a measurement of how clear the diamond is (I1 (worst), SI2, SI1, VS2, VS1, VVS2, VVS1, IF (best)) `nested bullet`    

* #### x `h4 heading, bullet`   
  * length in mm (0–10.74) `nested bullet`   

* #### y `h4 heading, bullet`   
  * width in mm (0–58.9) `nested bullet`   

* #### z `h4 heading, bullet`   
  * depth in mm (0–31.8) `nested bullet`   

* #### depth `h4 heading, bullet`   
  * `total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79)` `nested bullet, inline code block`    

* #### table `h4 heading, bullet`    
  * width of top of diamond relative to widest point (43–95) `nested bullet`

---


**For the next section you will need to load the following packages:**   

- knitr
- ggplot2
- kable
- pander

```{r, warning=FALSE}
# this is an example of a code chunk
#run manually by clicking the green arrow to the left if you need to load the library
library(knitr)
```

```{r, eval=FALSE, echo=FALSE}
# if you need to install the package this code chunk manually by clicking the green arrow tot he left
# you may be prompted to update packages. click `yes` to update.
install.packages("pander")
```

```{r, echo=FALSE}
library(pander)
```

### Next, you will practice working with code chunks!

```{r, echo=FALSE, eval=FALSE}
# This box is a code chunk example with 2 options:
# `echo=FALSE` - do not show the code chunck in knitted document
# `eval=FALSE` - do not run the code chunk when knitting a document
```



**3. Adding code chunks**

* On the tool bar, use the `Insert` button and selecting `R` or manually create on by typing a "code fence" 3 opening and closing backtics `with {r}  e.g. ```{r} at the beginning and closing with ``` ` to create a code chunk (see example above).    

* In between the "code fence" add and run the R code `summary(diamonds)` to get the summary of the data set: (run the code chunk to make sure it works)

  * **optional**- hide the code chunk in your knitted document by adding `echo=FALSE` in the code chunk options  

**Add code chunk below:**    

```{r, echo=FALSE}
summary(diamonds)
```





**4. Create a new code chunk to add a basic scatter plot to your document that shows the price as a function of carat weight:**

   * Use this code `plot(price ~ carat, data=diamonds)`
**Add code chunk below:**    

```{r}
plot(price ~ carat, data=diamonds)
```




**5. Create a basic R table that shows the count for the diamonds cut and clarity following these steps:**    

1. Create a code chunk
2. Hide the code chunk in the knitted document output using `echo=FALSE`
3. Use this code `table(diamonds$cut, diamonds$clarity)` to generate the table

**Add code chunk below:**   
```{r, echo=FALSE}
table(diamonds$cut, diamonds$clarity)
```




**Next, improve to formatting of this table by using the `knitr` package function `kable`:**

1. Use `kable` to imporve the look of the default table
2. Create a new code chunk with the option `echo=FALSE`
3. Use this code `kable(table(diamonds$cut, diamonds$clarity))` to gernerate the table
  * if you receive an error message, make sure the knitr library is enabled. run `library(knitr)` before you run the kable code.

**Add code chunk below:**   
```{r, echo=FALSE}
kable(table(diamonds$cut, diamonds$clarity))
```




**6. Create a publication style table using the pander package**

1. Install pander package using: `install.packages("pander")` run this code chunk if you need to install the pander package.
```{r, eval=FALSE, echo=FALSE}
install.packages("pander")
```

2. Load the pander package using: `library(pander)`   
```{r}
library(pander)
```

3. Run an simple anova (analysis of variance model) to model price as a function of carat.
    * Run this code to add the ANOVA to the variable model1:
```{r, eval=TRUE, results='hide'}
model1 <- aov(price ~ cut, data=diamonds)
summary(model1)
```
4. Gereate the table using the `pander package`.
    * Create a new code chunk with the option `echo=FALSE`
    * Generate the markdown output using pander
    * Use the code `pander(model1)`
    
**Add code chunk below:**   
```{r, echo=FALSE}
pander(model1)
```





**7. Add a lnked image to your document.**

  1. Use the image format `![image title](image link)` to embed a web linked image in the document.   
     * Use the sample image link to add an image to this document: `https://bit.ly/2ZQAIcg`   
     **Note: knitting a document with web linked images will only work when knitting to HTML.  Web linked images will produce a LaTex error when knitting to PDF.**

**Add Rmarkdown code below:**   

<!--![knitr workflow](https://bit.ly/2ZQAIcg)-->






**8. Knit your exercise worksheet to HTML or PDF output formats.**

  * Use the `knit` button in the tool bar to knit to the default setting
  * Use the small chevron next to `knit` to select optional output formats.
    


**Congratulations! You're on your way to making great documents using RStudio and Rmarkdown!** 




