---
title: "Software Setup"
output: html_document
---

<br>

To participate in either of these workshops, you will need a laptop with a user account that has software installation privileges to install the software described below. 
_Note that if you have separate user and admin accounts, you should run the installers as administrator (right-click on .exe file and select "Run as administrator" instead of double-clicking). Otherwise problems may occur later, for example when installing R packages._

In addition, you will need an up-to-date web browser.

To make best use of the workshop time please have the necessary programs and R packages installed prior to the start of the workshop. If you have any questions, click the _Contact_ link in the top right of this website to email the workshop organizer. 

Click the arrows to expand the section to show installation and setup instructions for each program. 

----


<details>
  <summary><font size="3">R (Both workshops)</font></summary>
    
**R** is a programming language that is especially powerful for data exploration, visualization, and statistical analysis. The project website is https://www.r-project.org. 

* **Windows** ([Video Tutorial](https://www.youtube.com/watch?v=q0PjTAylwoU)). Install R by downloading and running [this .exe file](https://cran.r-project.org/bin/windows/base/release.htm) from CRAN(https://cran.r-project.org/index.html).

* **macOS** ([Video Tutorial](https://www.youtube.com/watch?v=5-ly3kyxwEg)). Install R by downloading and running <a href="https://cran.r-project.org/bin/macosx/R-latest.pkg">this .pkg file</a> from <a href="https://cran.r-project.org/index.html">CRAN</a>. 

* **Linux** You can download the binary files for your distribution from CRAN(https://cran.r-project.org/index.html). Or you can use your package manager (e.g. for Debian/Ubuntu run `sudo apt-get install r-base` and for Fedora run `sudo dnf install R`). 


**Installation notes**:

* Default settings are OK. 
* Delete any desktop shortcut that was created
  
</details> 

<br>

<details>
  <summary>R Studio (Both workshops)</summary>

To interact with R, we use the _Integrated Development Environment (IDE)_ called **RStudio** (https://www.rstudio.com). The R Studio IDE is installed _after_ you have installed R. 

* Download R Studio (v 1.2+) from https://www.rstudio.com/products/rstudio/download/#download
    - Choose the download link that corresponds to your operating system. 

* Install R Studio
    * **Windows**: Double click on the downloaded file to run the installer program. 
    * **Mac**: Double click on the downloaded file, then drag the R Studio Icon into your Applications folder. 
    After you are done, eject the "Drive" that you downloaded by dragging it to your trash. 

* (Optional) Here is a short video shows you how to navigate R Studio: https://www.rstudio.com/products/RStudio/?wvideo=520zbd3tij

**Setting preferences in R Studio to retain sanity while debugging**

* Open R Studio and go to the file menu go to Tools then Global Options.
* Uncheck "Restore .RData into workspace at startup" 
* Where it says "Save workspace to .RData on exit:" Select "Never"
* Click apply then ok to close that window.
  
This will ensure that when you restart R you do not "carry forward" objects such as data sets that you were working on in a prior assignment. 

You should make a habit to completely shut down R studio when you are done working. 
Your open tabs will be saved, but your environment will be cleared. _This is a good thing._


</details> 

<br>

<details>
  <summary>R Packages (Both workshops)</summary>

R is considered an **Open Source** software program. That means many (thousands) of people contribute to the software. They do this by writing commands (called functions) to make a particular analysis easier, or to make a graphic prettier.

When you download R, you get access to a lot of functions that we will use. However these other _user-written_ packages add so much good stuff that it really is the backbone of the customizability and functionality that makes R so powerful of a language. 

> You only have to install packages once per R installation.  

To get started type in the **console** the following command to install the `rmarkdown` package. 

```r
install.packages("rmarkdown")
```

When the console returns to showing a `>` (and the stop sign in the top right corner of the Console window goes away), R is done doing what you asked it to do and ready for you to give it another command. You should get a message that looks similar to the message below. 

```
The downloaded binary packages are in
	C:\Users\Robin\AppData\Local\Temp\Rtmpi8NAym\downloaded_packages
```

Some packages use functions that live in other packages. These are called dependencies and will be automatically installed when you tell R to install a new package. 99% of the time you will not have to install them manually on your own. 


Now that you're a package installing pro, go ahead and install the following packages that pertain to your selected workshop. 


`r emo::ji("book")` **Workshop 12: Reproducible Reports using R Markdown**

* knitr
* rmarkdown
* ggplot2 
* kable
* pander


`r emo::ji("globe")` **Workshop 13: Website building**

* knitr
* rmarkdown
* devtools 
* blogdown 


<span style="color:red">R is case sensitive. If you see a message such as the one below, check for a typo.</span>

```
Warning in install.packages :
  package ‘ggplot’ is not available (for R version 3.5.1)
```

The correct package name is `ggplot2`, not `ggplot`.

</details>


<br>

<details>
  <summary>LaTeX (Reproducible Research workshop only)</summary>

To create a PDF directly from your source code you will have to install a lightweight version of $LaTeX$ (a document processing program). This allows you to make professional looking reports. 

Run the following code lines one at a time, waiting for the first one to finish before starting the next. This process can take up to 10 minutes. 
```{r, eval=FALSE}
install.packages("tinytex")
tinytex::install_tinytex()
```

If at any time it asks you to update or install any other packages, say yes to all of them. 

</details> 


<br>

<details>
  <summary>Git & GitHub (Website building workshop only)</summary>
  
Git is a version control system that lets you track who made changes to what when and has options for easily updating a shared or public version of your code on https://github.com. 

**Instructions to install Git can be found at** https://www.atlassian.com/git/tutorials/install-git

  
To share your code and publish your website you will also need a GitHub account at https://github.com. Basic GitHub accounts are free. We encourage you to create a GitHub account prior to the workshop if you don't have one already. Please consider what personal information you'd like to reveal. For example, you may want to review [these instructions for keeping your email address private](https://help.github.com/articles/keeping-your-email-address-private/) provided at GitHub.

> Note: This workshop is not an lesson on git. We will use git to do a few things and we will interact with GitHub, but we will not go into details about version control. 

</details> 


<br>

<details>
  <summary>Hugo (Website building workshop only)</summary>
  
Hugo (https://gohugo.io) is a static site generator, that is used by the R package `blogdown` to create dynamic websites. `blogdown` has a helper function to do this installation for us. 

```{r, eval=FALSE}
blogdown::install_hugo()
```

</details> 
